---
title: "Evaluations of positive and negative stimuli using the Affective Misattribution Procedure (AMP) and self-reports"
subtitle: "Data processing"
author: "Template: Ian Hussey; content: [Student name]"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r, include=FALSE}

knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)

```

# Dependencies

```{r}

library(tidyverse)
library(janitor) # for clean_names()
library(stringr)
library(openxlsx)

```

# Get data

```{r}

# demographics
data_demographics_raw <- read_csv("../data/raw/data_demographics_raw.csv") |>
  janitor::clean_names()

# data_demographics_raw_messy <- read_csv("../data/raw/data_demographics_raw_messy.csv", skip = 2) |>
#   janitor::clean_names()

# self report measure
data_selfreport_raw <- read_csv("../data/raw/data_selfreport_raw.csv") |>
  janitor::clean_names()

# affect attribution procedure
data_amp_raw <- read_csv("../data/raw/data_amp_raw.csv") |>
  janitor::clean_names()

```

# Demographics

```{r}

dat_age_gender <- data_demographics_raw |>
  select(subject, date, time, trialcode, response) |>
  pivot_wider(names_from = trialcode,
              values_from = response) |>
  mutate(gender = tolower(gender),
         gender = stringr::str_remove_all(gender, regex("\\W+")), # regex is both very useful and awful to write
         gender = case_when(gender == "female" ~ gender,
                            gender == "male" ~ gender,
                            gender == "nonbinary" ~ gender,
                            gender == "woman" ~ "female",
                            gender == "man" ~ "male",
                            TRUE ~ "other/missing/error"),
         age = case_when(str_detect(age, "^[0-9]+$") ~ age, # if the value is only numbers, keep it. 
                         TRUE ~ "other/missing/error")) 

```

# Exclusions / data quality

## AMP

```{r}

data_amp_performance_criteria <- data_amp_raw |> 
  filter(blockcode != "practice", 
         trialcode != "instructions") |> 
  mutate(latency_prob = if_else(latency < 100, TRUE, FALSE)) |> 
  group_by(subject) |> 
  summarize(proportion_fast_trials_amp = mean(latency_prob)) |>
  mutate(exclude_amp_performance = ifelse(proportion_fast_trials_amp > 0.10, "exclude", "include"))

# determine modal number of trials
data_amp_completeness <- data_amp_raw |>
  filter(blockcode != "practice",
         trialcode != "instructions") |>
  group_by(subject) |>
  count() |>
  ungroup() |>
  mutate(exclude_amp_completeness = ifelse(n == 72, "include", "exclude")) |>
  select(-n)

# data_amp_completeness |>
#   count(n)

```

- One participant with 8 trials appears to be a partial completion (check raw data?)
- One participant with 144 trials appears to be a repeat participant. I've chosen to exclude them entirely, but you could also have a more elaborate strategy where you retain only their first completion.

# Self-reports

```{r}

# trial level data
data_selfreport_trial_level <- data_selfreport_raw |>
  select(subject, trialcode, response) |>
  filter(trialcode %in% c("like", "prefer", "positive")) |>
  rename(item = trialcode) |>
  filter(response != "Ctrl+'B'") |>
  mutate(response = as.numeric(response))

# mean scored
data_selfreport_mean_score <- data_selfreport_trial_level |>
  group_by(subject) |>
  summarize(mean_evaluation = mean(response, na.rm = TRUE))

# combined
data_selfreport_scored <- 
  full_join(data_selfreport_trial_level |>
              pivot_wider(names_from = "item",
                          values_from = "response"),
            data_selfreport_mean_score,
            by = "subject")

```

# Affect Misattribution Procedure

TODO extract evaluations on the AMP test blocks and convert to an overall bias score

```{r}

data_amp_score_congruence <- data_amp_raw |> 
  select(subject, evaluative_response = correct, trialcode, blockcode) |> 
  filter(blockcode != "practice", 
         trialcode != "instructions") |> 
  mutate(trialcode = case_when(trialcode == "prime_positive" ~ 1, 
                               trialcode == "prime_negative" ~ 0,
                               TRUE ~ NA),
         prime_congruence = ifelse(trialcode == evaluative_response, 1, 0)) 

# sanity check 1: if you consider all the combiantions of factor levels of trialcode, evaluative_response, and prime congruence, there should be only 4:
data_amp_score_congruence |>
  count(trialcode, evaluative_response, prime_congruence)

data_amp_score_congruence |>
  count(trialcode, evaluative_response, prime_congruence) |>
  nrow() == 4

# calculate AMP score 
data_amp_score <- data_amp_score_congruence |> 
  group_by(subject) |> 
  summarize(AMP_score = mean(prime_congruence, na.rm = TRUE)) |> 
  select(subject, AMP_score)

# sanity check 2: check if AMP_score is numeric 
is.numeric(data_amp_score$AMP_score)

# sanity check 3: check if AMP_score is bounded [0,1]
data_amp_score |> 
  mutate(bounded_correctly = between(AMP_score, left = 0, right = 1)) |>
  filter(bounded_correctly != TRUE) |>
  nrow() == 0

```

# Combine

```{r}

# combine all dfs created in the previous chunks
data_processed_temp <- dat_age_gender |>
  full_join(data_selfreport_scored, by = "subject") |> 
  full_join(data_amp_score, by = "subject") |> 
  full_join(data_amp_performance_criteria, by = "subject") |>
  full_join(data_amp_completeness, by = "subject")

# flag all subjects with more than one row in the wide-format data. these should be excluded in the analysis.
# a more elaborate approach would be to track down the individual dupicate cases and determine which of the mulitiple cases should be retained. 
data_processed_duplicates <- data_processed_temp |>
  count(subject) |>
  mutate(exclude_duplicate_data = if_else(n > 1, "exclude", "include")) |>
  select(-n)

# join in the duplicates df
data_processed_before_exclusions <- data_processed_temp |>
  full_join(data_processed_duplicates, by = "subject")

```

# Define master exclusions

```{r}

# create a master exclude_participant variable
data_processed <- data_processed_before_exclusions |>
  mutate(exclude_participant = case_when(tolower(age) == "test" ~ "exclude",
                                         tolower(gender) == "test" ~ "exclude",
                                         is.na(mean_evaluation) ~ "exclude",
                                         # in this case we will exclude participants with missing demographics data or outcomes measures data. 
                                         # Note that "list-wise exclusions" like this aren't always justified, as missingness often isn't at random. 
                                         # How to treat missing data is a  whole area of work in itself, which we wont cover here.
                                         is.na(age) ~ "exclude", 
                                         is.na(gender) ~ "exclude",
                                         exclude_amp_performance == "exclude" ~ "exclude",
                                         exclude_duplicate_data == "exclude" ~ "exclude",
                                         exclude_amp_completeness == "exclude" ~ "exclude", 
                                         TRUE ~ "include"))

```

# Write to disk

```{r}

# in case this dir doesn't exist, create it
dir.create("../data/processed/")

# save data to disk in that dir
write_csv(data_processed, "../data/processed/data_processed.csv")

```

# Create codebook template for the processed data

If it has not already been created, this code write the codebook template to disk. 

\TODO The template should then be filled in manually with descriptions of each variable so that someone else could understand what these variables represent. 

```{r}

if(!file.exists("../data/processed/data_processed_codebook.xlsx")){
  # convert the column names to a df
  codebook_template <- data.frame(variable = colnames(data_processed)) |>
    mutate(explanation = NA)
  # write to disk as an excel file
  write.xlsx(codebook_template, file = "../data/processed/data_processed_codebook.xlsx")
}

```

Note that there are other ways of automatically creating more elaborate codebooks from your datasets. These often contain information about min/max/mean/SD, distribution, etc. For example:

- Ruben Arslan's {codebook}
  - [R package with How-Tos](https://rubenarslan.github.io/codebook/)
  - [Tutorial](https://rubenarslan.github.io/codebook/articles/codebook_tutorial.html)
  - [Article](https://journals.sagepub.com/doi/full/10.1177/2515245919838783)
- Petersen & Ekstrøm's {dataReporter}
  - [Article](https://www.jstatsoft.org/article/view/v090i06)
  - [Blog](https://sandsynligvis.dk/2017/08/21/datamaid-your-personal-assistant-for-cleaning-up-the-data-cleaning-process/)
  - [R package](https://cran.r-project.org/web/packages/dataReporter/index.html)

# Session info

```{r}

sessionInfo()

```


